---
Title: "Reproducible Research: Peer Assessment 1"
Output: html_document
keep_md: true
---

#Reproducible Research: Peer Assessment 1

### 1. Loading and preprocessing the data
```{r, echo=TRUE}
file <- unzip("activity.zip")
dfActivity <- read.csv(file, stringsAsFactors = FALSE)
dfActivity$date <- as.Date(dfActivity$date)
str(dfActivity)  
```  
  
### 2. What is mean total number of steps taken per day?
```{r, echo=TRUE}
dfActivity_rm<-dfActivity[which(!is.na(dfActivity$steps)),]
stepsPerDay<-tapply(dfActivity_rm$steps, dfActivity_rm$date, sum)

hist(stepsPerDay, 10,
      main = "Histogram of Daily Steps Taken from Oct to Nov 2012",
      xlab='Total Steps Each Day',
      ylab='Frequency',
      ylim=c(0, 20))

cat("Mean:", mean(stepsPerDay))  
cat("Median:", median(stepsPerDay))

```  

### 3. What is the average daily activity pattern?
```{r, echo=TRUE}
avgDailyActivity<-tapply(dfActivity_rm$steps, dfActivity_rm$interval, mean)
plot(y = avgDailyActivity, 
     x = names(avgDailyActivity), 
     type = "l", 
     xlab = "5-Minute-Interval", 
     main = "Average Daily Activity Pattern", 
     ylab = "Average number of steps")

avgDailyActivity[avgDailyActivity==max(avgDailyActivity)]
```
The 5-minute interval that, on average, contains the maximum number of steps is at interval 835.  
It recorded a total average of 206.17 steps.  
  

### 4. Imputing missing values
Current situation: The total number of missing values is 2,304 of the total 17,568 observations. All of the missing values fall under the 'Steps' variable.  
The strategy for imputing missing data: Replace missing values (na) with the mean of the 5-minute interval.

```{r, echo=TRUE}
sum(is.na(dfActivity))
sum(is.na(dfActivity$steps))
dfActivity_new <- dfActivity
dfActivity_new[which(is.na(dfActivity_new$steps)),1] <-
        avgDailyActivity[as.character(dfActivity_new[which(is.na(dfActivity_new$steps)),3])]
sum(is.na(dfActivity_new))

stepsPerDay_new<-tapply(dfActivity_new$steps, dfActivity_new$date, sum)

par(mfrow=c(1,2))
hist(stepsPerDay, 10, 
     main = "Original - Average Daily Steps", 
     xlab = "Steps", 
     xlim =c(0, 25000),
     ylim =c(0, 30))
abline(v = median(stepsPerDay), col = 4, lwd = 4)

hist(stepsPerDay_new, 10, 
      main = "Imputed - Average Daily Steps", 
      xlab = "Steps",
      xlim =c(0, 25000),
      ylim =c(0, 30))
abline(v = median(stepsPerDay_new), col = 4, lwd = 4)

mean(stepsPerDay_new)-mean(stepsPerDay)
median(stepsPerDay_new)-median(stepsPerDay)
```
The impact of imputing missing data with the mean of 5-minute interval. See results above:   
 * For the mean, there is no impact  
 * For the median, there is minimal impact where the median is increased by 1.189    

  

### 5. Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE }
dayType <- function(dates) 
{
  f <- function(date) 
    {
      if (weekdays(date) %in% c("Saturday", "Sunday")) 
        {
          "weekend"
        }
      else 
        {
        "weekday"
        }
    }
  sapply(dates, f)
}

dfActivity_new$dayType <- as.factor(dayType(dfActivity_new$date))

library(lattice)
dfDisplay <- aggregate(steps ~ interval + dayType, dfActivity_new, mean)
xyplot(steps ~ interval | dayType, 
       data=dfDisplay, 
       ylim = c(0,300), 
       layout=c(1,2), 
       type='l')
```

